---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse  = T,
  comment   = "#>",
  fig.path  = "man/figures/README-",
  out.width = "100%"
)
```
# multiarm

[![contributions welcome](https://img.shields.io/badge/contributions-welcome-brightgreen.svg?style=flat)](https://github.com/dwyl/esta/issues)
`r badger::badge_lifecycle("maturing", "blue")`
`r badger::badge_code_size("mjg211/multiarm")`
`r badger::badge_last_commit("mjg211/multiarm")`

## Description

`multiarm` provides functionality to assist with the design and analysis of
fixed-sample multi-arm clinical trials utilising one of several supported
multiple comparison corrections. Available functions allow for sample size
determination (including for *A*-, *D*-, and *E*-optimal designs), trial
simulation, analytical operating characteristic calculation (including the
conjunctive power, disjunctive power, family-wise error rate, and false
discovery rate), and the production of several informative plots. An R Shiny
graphical user interface is also provided to ease design determination.

## Getting started

You can install the latest development version of multiarm from
[Github](https://github.com/) with:

``` r
devtools::install_github("mjg211/multiarm/")
```

An introductory example of how to make use of the package's core functionality
can be found below. More detailed support is available in the package vignette,
which can be accessed with `vignette("multiarm")`. For further help, please
contact Michael Grayling at michael.grayling@newcastle.ac.uk.

## Details

Eight key functions are provided: `an_ma()`, `build_ma()`, `des_int_ma()`,
`des_ma()`, `gui_ma()`, `opchar_ma()`, `plot.multiarm_des_ma()`, and
`sim_ma()`. Outputs from all but `gui_ma()` and `plot.multiarm_des_ma()` are
also supported by S3 `print()` and `summary()` generics.

- `an_ma()`: Analyses summary statistics from a supplied fixed-sample multi-arm
clinical trial design, in order to determine which null hypotheses to reject.
- `build_ma()`: Builds a fixed-sample multi-arm clinical trial design object,
like those returned by `des_ma()` and `des_int_ma()`. For use when a specific
design is of interest.
- `des_int_ma()`: Determines the *A*-, *D*-, and *E*-optimal allocation of a set
of patients in a fixed-sample multi-arm clinical trial.
- `des_ma()`: Determines the sample size required a fixed-sample multi-arm
clinical trial when one of a variety of multiple comparison procedures is used.
- `gui_ma()`: Provides a graphical user interface to design determination.
- `opchar_ma()`: Analytically determines the operating characteristics (power,
family-wise error rates, etc.) of a supplied fixed-sample multi-arm clinical
trial design.
- `plot.multiarm_des_ma()`: Produces informative plots (power, false discovery
rate curves, etc.) relating to a supplied fixed-sample multi-arm clinical trial design.
- `sim_ma()`: Empirically estimates the operating characteristics (power,
family-wise error rates, etc.) of a supplied fixed-sample multi-arm clinical
trial design, via simulation.
 
## Example

Typically, `des_ma()` is used first to identify a design for the trial
parameters of interest. For example, consider designing a trial for

- three experimental treatment arms (see `K`);
- desiring a familywise error-rate of at most 5\%, controlling using Dunnett's
correction (see `alpha` and `correction`);
- desiring marginal power to reject each null hypothesis of 80\% for a
clinically relevant difference of 1 (see `beta`, `delta1`, and `power`);
- assuming the standard deviation of all responses is 1 (see `sigma`);
- allocating patients equally to each arm (see `ratio`).

To compute the implied design, we would run

```{r}
des <- multiarm::des_ma(K          = 3,
                        alpha      = 0.05,
                        beta       = 0.2,
                        delta1     = 1,
                        sigma      = rep(1, 4),
                        ratio      = rep(1, 3),
                        correction = "dunnett",
                        power      = "marginal")
```

Then, the total required sample size is

```{r}
des$N
```

In addition, the operating characteristics under the *global null*,
*global alternative*, and each of the *least favourable configurations*, can be
accessed with

```{r}
des$opchar
```

Useful plots can then be produced with `plot.multiarm_des_ma()` as follows

```{r dpi = 600}
multiarm:::plot.multiarm_des_ma(des)
```
