---
title: "Fixed-sample multi-arm clinical trial design"
date: "`r format(Sys.time(), '%d/%m/%Y')`"
params:
  K:            NA
  alpha:        NA
  beta:         NA
  delta1:       NA
  delta0:       NA
  sigma:        NA
  ratio_type:   NA
  ratio_init:   NA
  ratio:        NA
  correction:   NA
  power:        NA
  integer:      NA
  large_N:      NA
  small_n:      NA
  opchar:       NA
  pi:           NA
  piO:          NA
  plots:        NA
  equal:        NA
  shifted:      NA
  data:         NA
---

```{r, include = F}
knitr::opts_chunk$set(echo    = F,
                      results = "asis")
```

# Design setting

The trial will be designed to compare $K$ experimental treatments to a shared
control arm. Response $X_{ik}$, from patient $i=1,\dots,n_k$ in arm
$k=0,\dots,K$, will be assumed to be distributed as
$X_{ik} \sim N(\mu_k,\sigma_k^2)$. Then, the hypotheses to be tested will be:
$$ H_k : \tau_k = \mu_k - \mu_0 \le 0,\ k=1,\dots,K.$$
The *global null hypothesis*, $H_G$, will be:
$$ \tau_1 = \cdots = \tau_K = 0. $$
The *global alternative hypothesis*, $H_A$, will be:
$$ \tau_1 = \cdots = \tau_K = \delta_1. $$
The *least favourable configuration* for experimental arm $k$, $LFC_k$, will be:
$$ \tau_k = \delta_1,\ \tau_1 = \cdots = \tau_{k-1} = \tau_{k+1} = \cdots = \tau_K = \delta_0. $$
Here, $\delta_1$ and $\delta_0$ are *interesting* and *uninteresting* treatment
effects respectively.

# Inputs

```{r, echo = F, results = "asis"}
  K          <- params$K
  alpha      <- params$alpha
  beta       <- params$beta
  delta1     <- params$delta1
  delta0     <- params$delta0
  sigma      <- params$sigma
  ratio_type <- params$ratio_type
  ratio_init <- params$ratio_init
  ratio      <- params$ratio
  correction <- params$correction
  power      <- params$power
  integer    <- params$integer
  N          <- params$large_N
  n          <- params$small_n
  opchar     <- params$opchar
  pi         <- params$pi
  piO        <- params$piO
  plots      <- params$plots
  equal      <- params$equal
  shifted    <- params$shifted
  delta      <- params$delta
```
  
```{r, echo = F, results = "asis"}
if (correction == "bonferroni") {
  correction_text     <- "Bonferroni's correction"
} else if (correction == "dunnett") {
  correction_text     <- "Dunnett's correction"
} else if (correction == "none") {
  correction_text     <- "no multiple comparison correction"
} else if (correction == "sidak") {
  correction_text     <- "Sidak's correction"
} else if (correction == "benjamini_hochberg") {
  correction_text     <- "the Benjamini-Hochberg correction"
} else if (correction == "hochberg") {
  correction_text     <- "Hochberg's correction"
} else if (correction == "holm") {
  correction_text     <- "Holm's correction"
} else if (correction == "step_down_dunnett") {
  correction_text     <- "the step-down version of Dunnett's correction"
}

if (power == "marginal") {
  power_type_text     <- "marginal power for each null hypothesis"
  power_scenario_text <-
    "each of their respective least favourable configurations"
} else {
  power_scenario_text <- "the global alternative hypothesis."
  if (power == "conjunctive") {
    power_type_text   <- "conjunctive power"
  } else if (power == "disjunctive") {
    power_type_text   <- "disjunctive power"
  }
}

if (ratio_type == "equal_all") {
  ratio_text          <- "the same as the control arm"
} else if (ratio_type == "equal_exp") {
  ratio_text          <- paste0(ratio_init[1], " times that of the control arm")
} else if (ratio_type == "unequal") {
  ratio_text          <- paste0(paste(ratio_init[1:K], collapse = ", "),
                                " times that of the control arm.")
} else if (ratio_type == "root_K") {
  ratio_text          <- paste0(round(1/sqrt(K), 3),
                                " times that of the control arm.")
} else {
  ratio_text          <- paste0("chosen for ", ratio_type, "-optimality")
}

if (integer) {
  integer_text        <- "will"
} else {
  integer_text        <- "will not"
}

if (plots) {
  plot_text           <- "will"
} else {
  plot_text           <- "will not"
}
```

The following choices were made:

- $K = \boldsymbol{`r K`}$ experimental treatments will be included in the
trial.
- A significance level of $\alpha = \boldsymbol{`r alpha`}$ will be used, in
combination with __`r correction_text`__.
- The standard deviations of the responses will be assumed to be: $(\sigma_0,\dots,\sigma_K) = \boldsymbol{(`r paste(round(sigma, 3), collapse = ", ")`)}$.
- The __`r power_type_text`__ will be controlled to level
$1-\beta = \boldsymbol{`r 1 - beta`}$ under __`r power_scenario_text`__.
- The interesting and uninteresting treatment effects will be:
$\delta_1 = \boldsymbol{`r round(delta1, 3)`}$ and
$\delta_0 = \boldsymbol{`r round(delta0, 3)`}$ respectively.
- The target allocation to each of the experimental arms will be:
__`r ratio_text`__.
- The sample size in each arm __`r integer_text`__ be required to be an integer.
- Plots __`r plot_text`__ be produced.

# Outputs

```{r, echo = F, results = "asis"}
FWER                  <- round(opchar$FWER[1], 3)
if (correction %in% c("benjamini_hochberg", "hochberg", "holm",
                      "step_down_dunnett")) {
  punc_text_3         <- "s"
  pi_text             <- paste0("(", paste(round(piO, 3), collapse = ", "), ")")
} else {
  punc_text_3         <- ""
  pi_text             <- round(pi, 3)
}

if (power == "marginal") {
  power_type_text     <- "minimum marginal power"
  power_value         <-
    round(min(diag(as.matrix(opchar[-(1:2), (K + 3):(2*K + 2)]))), 3)
} else {
  power_scenario_text <- "the global alternative hypothesis."
  if (power == "conjunctive") {
    power_type_text   <- "conjunctive power"
    power_value       <- round(opchar$Pcon[2], 3)
  } else if (power == "disjunctive") {
    power_type_text   <- "disjunctive power"
    power_value       <- round(opchar$Pdis[2], 3)
  }
}
```

- The total required sample size is: $N = \boldsymbol{`r round(N, 3)`}$.
- The required sample size in each arm is: $(n_0,\dots,n_K) = \boldsymbol{(`r paste(round(n, 3), collapse = ", ")`)}$.
- Therefore, the realised allocation ratios to the experimental arms are:
$(r_1,\dots,r_K) = \boldsymbol{(`r paste(round(ratio, 3), collapse = ", ")`)}$.
- The maximum familywise error-rate is: $\boldsymbol{`r FWER`}$.
- The __`r power_type_text`__ is: $\boldsymbol{`r power_value`}$.
- The following critical threshold`r punc_text_3` should be used with the chosen
multiple comparison correction: $\boldsymbol{`r pi_text`}$.

```{r, echo = F, results = "asis"}
library(knitr)
kable(params$data)
```

# Plots

```{r, echo = F, results = "asis", dpi = 600}
if (params$plots) {
  print(params$equal)
  cat("\n \n \n \n \n \n")
  print(params$shifted)
} else {
  cat("Not chosen for output: select the box next to 'Plot power curves' to",
      "add plots to the report.")
}
```
